A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-04-09 manuscript.Rmd
added new intro paragraph

The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses. Next, one needs to make sure that all files relevant for the analysis can be moved to another computer. To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder. A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.

@@

## Reproducing an Analysis

We provide a reproducible analysis as a working example via [GitHub](https://github.com/aaronpeikert/workflow-showcase) as a testbed for users. The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'. The analysis demonstrates a successful implementation of our workflow including downloads of external data, comparison of their integrity using a checksum, and the running of a confirmatory factor analysis on the first few items using `lavaan` [@lavaan2012]. Once all mentioned tools are installed on a computer, the following few steps are necessary to reproduce our analysis:

```{bash, eval=FALSE}
git clone https://github.com/aaronpeikert/workflow-showcase.git
cd workflow-showcase
make build
make all DOCKER=TRUE
```

>>>>

# The Reproducible Workflow in a nutshell

![](Images/placeholder.png)
Figure XXX gives an overview of how the four components of our workflow interface to ensure computational reproducibility. We begin with a minimal description of the roles of the four components. In the remainder of this tutorial, we will further detail each of the four components of our workflow. The first component is version control. Version control manages  changes to files (e.g., data and code) over time so that you can access specific versions later. Version control offers snapshots of your workflow at different time points identified by a unique identifier. The different frames in the z-axis of the image illustrate the different versions of the workflow. How different parts of a analysis and a report of it relate and in what order they need to be executed is monitored using dependency management. The arrows visualize dependencies, such as an analysis depending on the availability of a particular data file. Third, all computer code (such as a statistical analysis in R) is executed in a virtual environment that guarantuees exact reproduction of results independent of the host operating system, R version, and installed package versions. Finally, literate programming interweaves human-readable code and automatically produced results (such as p values, effect sizes, confidence intervals) to eliminate inconsistency errors such as those arising from careless copy and paste actions.

@@

Finally, we conclude with some notes on project organization, which we think makes migrating projects to a reproducible workflow easier. The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses. Next, one needs to make sure that all files relevant for the analysis can be moved to another computer. To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder. A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.

## Use Case: Reproducing an Analysis

We provide a reproducible analysis as a working example via [GitHub](https://github.com/aaronpeikert/workflow-showcase). We encourage interested readers to try to reproduce this example as a practical exercise. The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'. The analysis demonstrates a complete implementation of our workflow including downloads of external data, comparison of their integrity using a checksum, and  a confirmatory factor analysis on the first few items using the R package `lavaan` [@lavaan2012]. Once all required tools are installed on a computer, the following four command-line commands are sufficient to reproduce our demo analysis:

```{bash, eval=FALSE}
# (1) download the complete repository containing the workflow to your computer
git clone https://github.com/aaronpeikert/workflow-showcase.git
# (2) enter the directory containing the cloned workflow
cd workflow-showcase
# (3) building the docker container
make build
# (4) running the analysis and producing the final PDF inside the container
make all DOCKER=TRUE
```
